---
title: "A3: SPL Library Checkouts"
author: "Natalie Olson"
date: "02/22/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
# load and install ALL necessary libraries
library(dplyr)
library(tidyverse)
library(readr)
library(ggplot2)
library("ggalt")
library("lubridate")
library("viridis")
library("plotly")
library("scales")
# load library dataframe
library_df <- read.csv("./2017-2023-10-Checkouts-SPL-Data.csv")
leigh_bardugo_all_df <- library_df %>% 
  filter(Creator %in% c("Leigh Bardugo", "Bardugo, Leigh")) %>% 
  filter(!CheckoutYear == 2023)
leigh_bardugo_all_df$Creator <- tolower(leigh_bardugo_all_df$Creator)
leigh_bardugo_all_df$Title <- tolower(leigh_bardugo_all_df$Title)
leigh_bardugo_all_df$Title[str_detect(leigh_bardugo_all_df$Title, "shadow and bone")] <- "Shadow and Bone"
leigh_bardugo_all_df$Title[str_detect(leigh_bardugo_all_df$Title, "siege and storm")] <- "Siege and Storm"
leigh_bardugo_all_df$Title[str_detect(leigh_bardugo_all_df$Title, "ruin and rising")] <- "Ruin and Rising"
leigh_bardugo_all_df$Title[str_detect(leigh_bardugo_all_df$Title, "six of crows")] <- "Six of Crows"
leigh_bardugo_all_df$Title[str_detect(leigh_bardugo_all_df$Title, "crooked kingdom")] <- "Crooked Kingdom"
leigh_bardugo_all_df$Title[str_detect(leigh_bardugo_all_df$Title, "king of scars")] <- "King of Scars"
leigh_bardugo_all_df$Title[str_detect(leigh_bardugo_all_df$Title, "rule of wolves")] <- "Rule of Wolves"
grishaverse_df <- leigh_bardugo_all_df %>% 
  filter(Title %in% c("Shadow and Bone", "Siege and Storm", "Ruin and Rising",
                             "Six of Crows", "Crooked Kingdom", "King of Scars", "Rule of Wolves"))
summary_info <- list()
# source all R files
source('chart1.R')
source('chart2.R')
source('chart3.R')
source('chart3.R')
source('summary.R')
```
### Introduction

With the increasing integration of technology in people's lives, fewer people look to books as a source of entertainment. When the COVID pandemic hit, people finally turned back to books after a steady decline in the number of readers that had been lowering since the beginning of the twenty-first century. With everyone stuck at home with not much better to do, both book sales and library checkout numbers saw a massive increase, especially eBooks, considering they're more portable and accessible; anyone could check out an eBook from the library without even leaving their homes. In the young adult setting, the creation of BookTok brought many readers back to the hobby as well and introducing new readers through this online community. 

This report looks over Leigh Bardugo's *Grishaverse* novels and their trends in popularity over the years, as not only is she one of my personal favorite authors, but she is one of the most prominent young adult fantasy authors right now. This report hopes to discover:

- How did the pandemic affect library checkouts?
- Which of Leigh Bardugo's books are most commonly checked out? How do they compare to each other?
- Have Bardugo's books increased or decreased in popularity throughout the past few years (as inferred by library checkout data)?

This dataset is filtered to only include the seven *Grishaverse* novels (*Shadow and Bone, Siege and Storm, Ruin and Rising, Six of Crows, Crooked Kingdom, King of Scars*, and *Rule of Wolves*).

### Summary

From 2017 to 2022, Leigh Bardugo's *Grishaverse* novels have been checked out a total of **`r summary_info$total_checkouts`** times. The majority of people reading these novels choose to check out the **`r summary_info$grishaverse_material_type`** format rather than checking out physical books or audiobooks.
**`r summary_info$highest_checkouts`** is her most popular novel both historically and presently, and during this time period, it was checked out **`r summary_info$sab_checkouts`** times, while her least popular novel, **`r summary_info$lowest_checkouts`**, was only checked out **`r summary_info$row_checkouts`** times. Her most popular series, the **`r summary_info$highest_series`**, was checked out the most at a total of **`r summary_info$grisha_trilogy_checkouts`** times.

On average, Bardugo's *Grishaverse* novels are checked out **`r summary_info$avg_per_year`** times per year, and when this number is split between the seven novels, each *Grishaverse* novel should be checked out at least **`r summary_info$avg_per_novel`** times. We will observe that this will not be the case, as we can infer by **`r summary_info$lowest_checkouts`**'s total number of checkouts; **`r summary_info$highest_checkouts`** was checked out **`r summary_info$high_low_difference`** more times than **`r summary_info$lowest_checkouts`** during this time frame. On average, Bardugo's novels are being **checked out more** in current years, and it seems as thought they will continue to see a **steady growth** in popularity.


### The Dataset

Seattle Public Library collects their monthly checkout data and publishes it for the public. Some of the parameters in this data set include: title, publication date, checkout month and year, genre, creator, material type, and number of checkouts among other examples. The data is generated from a variety of sources, including: Overdrive, hoopla, Freegal, and RBDigital which provide electronic data for digital items while physical item checkout data was sourced from the Lebgrady artwork data archives from April 2005 to September 2016. Currently, physical item checkout data is collected from the Horizon ILS. This data set essentially acts as a monthly record of library checkouts, holding hundreds of thousands of checkouts over the years. In a sense, it acts as a marker of history from 2005 until the present.

At first glance, this data set is quite messy; certain parameters are formatted strangely, such as Publication Year which has around seven different potential formats that each mean different things. Many of the same books are referenced under different titles depending on the different material type. In addition, other fields have such a wide variety of different answers that it may be difficult to find trends that don't occur over such a vast margin, such as when sorting by subjects. There are gaps in the data here and there and plenty of NA values, and given the size of the data set, there is quite a lot to sort through. 

There are certain variables that would've been useful to track alongside what has already been collected. For example, we may be able to draw conclusions about a book's quality or success based on the number of checkouts, but there isn't any other statistic provided that may strengthen those assumptions such as an average rating from Goodreads or something along those lines. We have no way of knowing the demographics of the people who are checking out items, as all that is tracked is information about the items themselves, so we cannot draw any conclusions about people or culture in relation to checkouts without making inferences. Ultimately, this data set is just about library checkout data, so its scope only exists within that range.

### First Trends Over Time Chart

``````{r, echo = FALSE, message = FALSE, warning = FALSE, code = readLines("chart1.R")}
source("chart1.R")
``````

This line chart visualizes the total number of checkouts for each of the seven books across the *Grishaverse* series split into their respective series: the ***Dregs Series*** (*Six of Crows* and *Crooked Kingdom*), the ***Grisha Trilogy*** (*Shadow and Bone, Siege and Storm,* and *Ruin and Rising*), and the ***King of Scars Duology*** (*King of Scars* and *Rule of Wolves*).

The most popular series in terms of checkouts is the ***Grisha Trilogy***, followed by the ***Dregs Series***, and finally, the ***King of Scars Duology***. The ***Grisha Trilogy*** is the only series that increases steadily over time, while the other two series have had periods of an increase in checkouts or a decrease in checkouts. While the ***Grisha Trilogy*** has only increased in checkout numbers over the years, this number spiked significantly in 2021, which can be possibly explained by the release of the *Shadow and Bone* Netflix adaptation in April of 2021.

This dataframe only tracks data from 2017 onwards, but it looks as though the ***Dregs Series*** had a decrease in checkouts from 2017 to 2019 after the second book, *Crooked Kingdom*, was released in 2016. However, it did see an increase in checkouts around 2019 again around the release of the first *King of Scars* novel and spiked in 2021, around the time of the COVID pandemic, only to decrease again in 2022.

The ***King of Scars Duology*** has the lowest number of checkouts all around, although, in 2022, it started catching up to the ***Dregs Series***. Both of these series have a significantly lower number of checkouts in comparison to the ***Grisha Trilogy***. 


### Second Trends Over Time Chart

``````{r, echo = FALSE, message = FALSE, warning = FALSE, code = readLines("chart2.R")}
source("chart2.R")
source('summary.R')
``````

If you hover over the different bars, you can view the total number of checkouts for each novel each month.

This bar chart tracks total checkouts for each Grishaverse novel from January to December of 2021. Except for January and Feburary, where ***Six of Crows*** had the highest number of checkouts, ***Shadow and Bone*** is the most commonly checked out *Grishaverse* novel, while ***Rule of Wolves*** has the least number of checkouts by far. What's most striking about this chart is the steady increase from February to April until there is a sudden spike in numbers in **May**, with a total of **`r summary_info$total_may_checkouts`** checkouts between all novels compared to April's **`r summary_info$total_april_checkouts`** checkouts, which is **`r summary_info$may_april_difference`** more checkouts. 

The number of checkouts between May and December remain around the same area, with a slight decrease from May to September. **October** of 2021 saw the highest number of checkouts with **`r summary_info$total_oct_checkouts`** checkouts! 


### Third Chart

``````{r, echo = FALSE, message = FALSE, warning = FALSE, code = readLines("chart3.R")}
source("chart3.R")
``````

This pie chart displays the total number of checkouts for the *Grishaverse* novels as divided by respective novel. Leigh Bardugo's *Grishaverse* novels were checked out a total of **`r summary_info$total_checkouts`** times between 2017 and 2022. ***Shadow and Bone*** remains the most commonly checked out *Grishaverse* novel, while ***Rule of Wolves***, which was released in March of 2021, is by far the least popular, with only **95** total checkouts by the end of 2022. ***Six of Crows*** doesn't trail too far behind ***Shadow and Bone***, and both novels make up the majority of total checkouts. It's difficult to determine a pattern between the different novels, since while ***Shadow and Bone*** and ***Six of Crows*** are both the first novels in their series, ***King of Scars*** is one of the less popular *Grishaverse* novels to check out. This could be because it's the first novel of the third series or because it was published more recently. 

